---
title: "Lesson 8 - Diamonds Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Question 1

Create a histogram of diamond prices. Facet the histogram by diamond color and use cut to color the histogram bars.

The plot should look something like this: http://i.imgur.com/b5xyrOu.jpg

Note: In the link, a color palette of type 'qual' was used to color the histogram using `scale_fill_brewer(type = 'qual')`



```{r}

```



# Question 2

Create a scatterplot of diamond price vs. table and color the points by the cut of the diamond.

The plot should look something like this: http://i.imgur.com/rQF9jQr.jpg

Note: In the link, a color palette of type 'qual' was used to color the scatterplot using `scale_color_brewer(type = 'qual')`

```{r}

```




# Question 3

Use the graph you created from the previous exercise to see the answer. You do not need to run summaries.

(a) What is the typical table range for the majority of diamonds of `ideal` cut?

Answer:  _______ to ______

(b) What is the typical table range for the majority of diamonds of `premium` cut?

Answer:  _______ to ______




# Question 4

Create a scatterplot of diamond price vs volume (x * y * z) and color the points by the clarity of diamonds. Use scale on the y-axis to take the log10 of price. You should also omit the top 1% of diamond volumes from the plot.

Note: Volume is a very rough approximation of a diamond's actual volume.

The plot should look something like this: http://i.imgur.com/excUpea.jpg

Note: In the link, a color palette of type 'div' was used to color the scatterplot using `scale_color_brewer(type = 'div')`



# Question 5

Many interesting variables are derived from two or more others. For example, we might wonder how much of a person's network on a service like Facebook the user actively initiated. Two users with the same degree (or number of friends) might be very different if one initiated most of those connections on the service, while the other initiated very few. So it could be useful to consider this proportion of existing friendships that the user initiated. This might be a good predictor of how active a user is compared with their peers, or other traits, such as personality (i.e., is this person an extrovert?).

Your task is to create a new variable called 'prop_initiated' in the Pseudo-Facebook data set. The variable should contain the proportion of friendships that the user initiated.

```{r}
# DO NOT DELETE THIS NEXT LINE OF CODE
# ========================================================================
pf <- read.delim('/datasets/ud651/pseudo_facebook.tsv')


# ENTER YOUR CODE BELOW THIS LINE
# ========================================================================



```






# Question 6

Create a line graph of the median proportion of friendships initiated ('prop_initiated') vs. tenure and color the line segment by `year_joined.bucket`.

Recall, we created `year_joined.bucket` in Lesson 5 by first creating `year_joined` from the variable tenure. Then, we used the cut function on `year_joined` to create four bins or cohorts of users.

# (2004, 2009]
# (2009, 2011]
# (2011, 2012]
# (2012, 2014]

The plot should look something like this: http://i.imgur.com/vNjPtDh.jpg

OR this:http://i.imgur.com/IBN1ufQ.jpg

This assignment is not graded and will be marked as correct when you submit.

```{r}

```



# Question 7

Smooth the last plot you created of of `prop_initiated` vs tenure colored by `year_joined.bucket`. You can bin together ranges of tenure or add a smoother to the plot.

There won't be a solution image for this exercise. You will answer some questions about your plot in the next two exercises.

```{r}

```


# Question 8

On average, which group initiated the greatest proportion of its Facebook friendships? The plot with the smoother that you created in the last exercise can help you answer this question.

(a) People who joined prior to 2009
(b) People who joined 2009-2011
(c) People who joined 2011-2012
(d) People who joined after 2012

# Question 9

(a) For the group with the largest proportion of friendships initiated, what is the group's average (mean) proportion of friendships initiated? 

Answer: 



(b) Why do you think this group's proportion of friendships initiated is higher than the others?

Answer:


# Question 10

Create a scatter plot of the price/carat ratio of diamonds. The variable x should be assigned to cut. The points should be colored by diamond color, and the plot should be faceted by clarity.

The plot should look something like this: http://i.imgur.com/YzbWkHT.jpg.

Note: In the link, a color palette of type 'div' was used to color the histogram using `scale_color_brewer(type = 'div')`

```{r}

```


# Question 11

The Gapminder website contains over 500 data sets with information about the world's population. Your task is to continue the investigation you did at the end of Problem Set 4 or you can start fresh and choose a different data set from Gapminder.

If you’re feeling adventurous or want to try some data munging see if you can find a data set or scrape one from the web.

In your investigation, examine 3 or more variables and create 2-5 plots that make use of the techniques from Lesson 5.

You can find a link to the Gapminder website in the Instructor Notes.

Once you've completed your investigation, create a post in the discussions that includes:
       1. the variable(s) you investigated, your observations, and any summary statistics
       2. snippets of code that created the plots
       3. links to the images of your plots

```{r}

```


