---
title: "Lesson 4 - Diamonds Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading the dataset

Load the diamonds data set in RStudio. 

```{r}
library(ggplot2) #must load the ggplot package first
data(diamonds) #loads the diamonds data set since it comes with the ggplot package
```


You should notice that the diamonds data set will be listed as a 'Promise' in the workspace. This is a special object in R, and you need to run a command on the data to fully load the data set. Try running a summary:

```{r}
summary(diamonds)
```


You can learn more about the diamonds data set by executing the following code. The summary should appear in the help window (usually, it's the lower right console).

```{r}
?diamonds
```



Read the help file to determine the best color for the diamonds data set. When you run diamonds$color, you will notice that the levels of color are set as D < E < F < ...

This does not indicate that D is the worst color but rather an ordering to the colors.




# Question 1: 

## (a) How many observations are in the dataset?

```{r}

```


## (b) How many variables are in the data set?

```{r}

```


## (c) How many ordered factors are in the data set?

```{r}

```


## (d) What letter represents the best color for a diamond?

Answer: 











# Question #2:

Create a histogram of the price of all the diamonds in the diamond data set.

```{r}

```



# Question #3: 

Describe the shape and center of the price distribution. Include summary statistics like the mean and median.

```{r}

```


Answer: 





# Question #4: Diamond Counts

(a) How many diamonds cost less than $500?

```{r}

```


(b) How many diamonds cost less than $250?

```{r}

```


(c) How many diamonds cost $15,000 or more?

```{r}

```




# Question #5: 

Explore the largest peak in the price histogram you created earlier.

Try limiting the x-axis, altering the bin width, and setting different breaks on the x-axis.

You can save images by using the ggsave() command.
  ggsave() will save the last plot created.
  For example...
                  qplot(x = price, data = diamonds)
                  ggsave('priceHistogram.png')

ggsave currently recognises the extensions eps/ps, tex (pictex), pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).


```{r}

```


# Question #6: 

Break out the histogram of diamond prices by cut.

You should have five histograms in separate panels on your resulting plot.

```{r}

```


# Question #7: 

Answer the questions below. Check more than one option if there were ties. 

Cut options: Fair, Good, Very Good, Premium, and Ideal

(a) Which cut has the highest priced diamond?

```{r}

```


(b) Which cut has the lowest priced diamond?

```{r}

```


(c) Which cut has the lowest median price?

```{r}

```



# Question #8: 

In the two last exercises, we looked at the distribution for diamonds by cut.

Run the code below in R Studio to generate the histogram as a reminder.

```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut)
```


In the last exercise, we looked at the summary statistics for diamond price by cut. If we look at the output table, the the median and quartiles are reasonably close to each other.

```{r}
by(diamonds$price, diamonds$cut, summary)
```


This means the distributions should be somewhat similar, but the histograms we created don't show that.

The 'Fair' and 'Good' diamonds appear to have different distributions compared to the better cut diamonds. They seem somewhat uniform on the left with long tails on the right.

Let's look in to this more.

Look up the documentation for `facet_wrap` in R Studio.
Then, scroll back up and add a parameter to facet_wrap so that  the y-axis in the histograms is not fixed. You want the y-axis to be different for each histogram.

If you want a hint, check out the Instructor Notes.


```{r}

```




# Question #9: 

Create a histogram of price per carat and facet it by cut. You can make adjustments to the code from the previous exercise to get started.

Adjust the bin width and transform the scale of the x-axis using log10.

```{r}

```


# Question #10: 

Investigate the price of diamonds using box plots, numerical summaries, and one of the following categorical variables: cut, clarity, or color.

There won’t be a solution video for this exercise so go to the discussion thread for either BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT to share you thoughts and to see what other people found.

You can save images by using the ggsave() command.
ggsave() will save the last plot created.
For example...
                  qplot(x = price, data = diamonds)
                  ggsave('priceHistogram.png')

ggsave currently recognises the extensions eps/ps, tex (pictex), pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).


# Question #11: 

What is the price range for the middle 50% of diamonds with color D?

first quartile (25%): 
third quartile (75%): 

```{r}

```



What is the price range for the middle 50% of diamonds with color J?

first quartile (25%): 
third quartile (75%): 

```{r}

```


What is the IQR for diamonds with the best color?

```{r}

```


What is the IQR for diamonds with the best color?

```{r}

```




# Question #12: 

Investigate the price per carat of diamonds across the different colors of diamonds using boxplots.

Go to the discussions to share your thoughts and to discover
what other people found.

You can save images by using the ggsave() command.
ggsave() will save the last plot created.
For example...
                  qplot(x = price, data = diamonds)
                  ggsave('priceHistogram.png')

ggsave currently recognises the extensions eps/ps, tex (pictex), pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).


# Question #13: 

Investigate the weight of the diamonds (carat) using a frequency polygon. Use different bin widths to see how the frequency polygon changes. What carat size has a count greater than 2,000? Check all that apply.

Answer choices:

0.1
0.3
0.8
1.01
1.6
2.0
3.0
5.0






# Question #14: 

Data munging or data wrangling can take up much of a data scientist's or data analyst's time. There are two R packages that make these tasks easier in R: tidyr and dplyr.

tidyr - a package that reshapes the layout of your data

dplyr - a package that helps you transform tidy, tabular data

Review Data Wrangling in R (https://s3.amazonaws.com/udacity-hosted-downloads/ud651/DataWranglingWithR.pdf) to get a sense of how these packages allow you to manipulate data. You can use these packages to help you in the next programming task and in your future investigations.

There are some useful cheat sheets on RStudio's webpage (http://www.rstudio.com/resources/cheatsheets/). The Data Import and Data Transformation cheat sheets available at that site will be especially good references for working with tidyr and dplyr, respectively.



# Question #15: 

The Gapminder website contains over 500 data sets with information about the world's population. Your task is to download a data set of your choice and create 2-5 plots that make use of the techniques from Lesson 3.

You might use a simple histogram, a boxplot split over a categorical variable, or a frequency polygon. The choice is yours!

You can find a link to the Gapminder website in the Instructor Notes.

Once you've completed your investigation, create a post in the discussions that includes:
       1. any questions you answered, your observations, and summary statistics
       2. snippets of code that created the plots
       3. links to the images of your plots

You can save images by using the ggsave() command.
ggsave() will save the last plot created.
For example...
                  qplot(x = price, data = diamonds)
                  ggsave('priceHistogram.png')

ggsave currently recognises the extensions eps/ps, tex (pictex), pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).




# Question #16: 


Your task is to investigate the distribution of your friends' birth months and days.

Here some questions you could answer, and we hope you think of others.

# **********************************************************************

# How many people have the same birthday as you?
# (Reserve time with them or save money to buy them a gift!)

# Which month contains the most number of birthdays?

# How many birthdays are in each month?

# Which day of the year has the most number of birthdays?

# Do you have at least 365 friends that have birthdays on everyday
# of the year?

# **********************************************************************

You will need to do some data munging and additional research to complete this task. This task won't be easy, and you may encounter some unexpected challenges along the way. We hope you learn a lot from it though.

You can expect to spend 30 min or more on this task depending if you use the provided data or obtain your personal data. We also encourage you to use the lubridate package for working with dates. Read over the documentation in RStudio and search for examples online if you need help.

You'll need to export your Facebooks friends' birthdays to a csv file. You may need to create a calendar of your Facebook friends’ birthdays in a program like Outlook or Gmail and then export the calendar as a csv file.

Once you load the data into R Studio, you can use the strptime() function to extract the birth months and birth days. We recommend looking up the documentation for the function and finding examples online.

We've included some links in the Instructor Notes to help get you started.

Once you've completed your investigation, create a post in the discussions that includes:
   1. any questions you answered, your observations, and summary statistics
   2. snippets of code that created the plots
   3. links to the images of your plots

You can save images by using the ggsave() command.
ggsave() will save the last plot created.
For example...
                  qplot(x = price, data = diamonds)
                  ggsave('priceHistogram.png')

ggsave currently recognises the extensions eps/ps, tex (pictex), pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).


```{r}

```


